---
title: "MI/Stroke Inpatient"
format:
  docx:
    fig-cap-location: top
execute:
  echo: false
  message: false
  warning: false
lof: true
---

```{r}
# Load necessary packages ----
library(OmopSketch)
library(visOmopResults)
library(dplyr)
library(omopgenerics)
library(flextable)
library(ggplot2)
library(here)
library(CodelistGenerator)
library(stringr)
library(CohortCharacteristics)
library(purrr)
library(rlang)
library(tidyr)
library(scales)

cohort <- "mi"
# cohort <- "stroke"

if (cohort == "mi") {
  cohorts <- c("acute_mi_narrow", "acute_mi_narrow_stemi", "acute_mi_narrow_not_stemi")
  drugs <- c("antihypertensives", "beta_blockers", "anticoagulants", "antiplatelets", "lipid_lowering_drugs", "thrombolytics")
} else {
  cohorts <- c("stroke_broad")
  drugs <- c("anticoagulants", "antiplatelets", "lipid_lowering_drugs", "thrombolytics")
}

clean <- function(x) {
  x |>
    str_replace_all(pattern = "_", replacement = " ") |>
    str_to_title()
}

drugsTidy <- clean(drugs)

# Load results stored in the package ----
fileData <- file.path(getwd(), "data", "studyData.RData")
if (!file.exists(fileData)) {
  source(file.path(getwd(), "rawData", "preprocess.R"))
}

# uncomment to load the raw data
# rawData <- omopgenerics::importSummarisedResult(file.path(getwd(), "rawData"))

# load data
load(fileData)

# source functions
source(file.path(getwd(), "functions.R"))

data$summarise_treatments$variable_level <- str_to_title(data$summarise_treatments$variable_level)

# filter data
data <- data |>
  imap(\(x, nm) {
    if (!nm %in% c("summarise_omop_snapshot", "summarise_observation_period")) {
      if (inherits(x, "summarised_result")) {
        col <- c("cohort_name", "index_condition") |>
          keep(\(y) y %in% groupColumns(x))
        if (length(col) == 2) {
          x <- x |>
            filterGroup(cohort_name %in% cohorts | index_condition %in% cohorts)
        } else {
          x <- x |>
            filterGroup(.data[[col]] %in% cohorts)
          if (col == "index_condition") {
            x <- x |>
              filterStrata(drug %in% drugs)
          }
        }
      } else {
        x <- x |>
          filter(index_condition %in% cohorts & drug %in% drugs)
      }
      if (nm == "summarise_treatments") {
        x <- x |>
          filter(variable_level %in% drugsTidy)
      }
    }
    return(x)
  })

# Global options ----
knitr::opts_chunk$set(
  out.width  = "95%",  # figures occupy ~95% of document width
  out.height = "auto",
  dpi        = 320,    # ensure figure quality
  fig.width  = 6,      # default aspect ratio (can be overridden per-figure)
  fig.height = 3,
  results    = "asis"  # enable Markdown produced via cat() inside chunks
)

updateCohortName <- function(x, col, version = "long") {
  dic <- tribble(
    ~name, ~long, ~short,
    "stroke_broad", "Acute Ischaemic Stroke", "Stroke",
    "stroke_narrow", "Acute Ischaemic Stroke narrow", "Stroke narrow",
    "stroke_broad_with_af", "Acute Ischaemic Stroke with Atrial Fibrillation", "Stroke with AF",
    "stroke_narrow_with_af", "Acute Ischaemic Stroke narrow with Atrial Fibrillation", "Stroke narrow with AF",
    "acute_mi_broad", "Acute Myocardial Infarction broad", "MI broad",
    "acute_mi_narrow", "Acute Myocardial Infarction", "MI",
    "acute_mi_broad_stemi", "Acute Myocardial Infarction broad ST-Elevation Myocardial Infarction", "MI broad STEMI",
    "acute_mi_narrow_stemi", "Acute Myocardial Infarction ST-Elevation Myocardial Infarction", "MI STEMI",
    "acute_mi_broad_not_stemi", "Acute Myocardial Infarction broad Non-ST-elevation Myocardial Infarction", "MI broad NSTEMI",
    "acute_mi_narrow_not_stemi", "Acute Myocardial Infarction Non-ST-elevation Myocardial Infarction", "MI NSTEMI"
  )
  q <- paste0(
    ".data[['", col, "']] == '", dic$name, "' ~ '", dic[[version]], "'"
  )
  q <- paste0(c(q, ".default = .data[[col]]"), collapse = ", ")
  q <- paste0("case_when(", q, ")") |>
    parse_exprs() |>
    set_names(col)
  
  x |>
    filter(.data[[col]] %in% cohorts) |>
    mutate(!!!q)
}

setGlobalPlotOptions(type = "ggplot")
setGlobalTableOptions(type = "flextable")
```

```{r, echo=FALSE}
cdm <- omock::mockCdmFromDataset("delphi-100k")

comorbidities <- importCodelist(path = here("../study/Cohorts/comorbidities")) |>
  asCodelistWithDetails(cdm = cdm)
miTypes <- importCodelist(path = here("../study/Cohorts/miTypes")) |>
  asCodelistWithDetails(cdm = cdm)
names(miTypes) <- c("acute_mi_narrow_not_stemi", "acute_mi_narrow_stemi")
comorbidities <- c(comorbidities, miTypes)
drugsCodes <- importCodelist(path = here("../study/Cohorts/drugs"))[drugs] |>
  asCodelistWithDetails(cdm = cdm)
procedures <- importCodelist(path = here("../study/Cohorts/miProcedures")) |>
  asCodelistWithDetails(cdm = cdm)
```

:::{custom-style="Caption"}
**Table S1:** Data partners characteristics.
:::

```{r}
data$summarise_omop_snapshot |>
  filter(variable_name != "cdm_source" | estimate_name == "type") |>
  tableOmopSnapshot()
```

:::{custom-style="Caption"}
**Table S2:** Codelists for condition of interest.
:::

```{r}
comorbidities[cohorts] |>
  as_tibble() |>
  updateCohortName(col = "codelist_with_details_name") |>
  visTable(groupColumn = "codelist_with_details_name")
```

:::{custom-style="Caption"}
**Table S3:** Codelists for drugs
:::

```{r}
drugsCodes |>
  as_tibble() |>
  mutate(
    codelist_with_details_name = codelist_with_details_name |>
      str_replace_all("_", " ") |>
      str_to_title()
  ) |>
  # subset to ingredients
  inner_join(
    cdm$concept |>
      filter(concept_class_id == "Ingredient") |>
      select("concept_id"),
    by = "concept_id"
  ) |>
  visTable(groupColumn = "codelist_with_details_name")
```

:::{custom-style="Caption"}
**Table S4:** Codelists for Coronary Artery Bypass Graft and Percutaneous Coronary Intervention.
:::

```{r, echo=FALSE}
procedures |>
  as_tibble() |>
  mutate(
    codelist_with_details_name = codelist_with_details_name |>
      str_replace_all("_", " ") |>
      str_to_title()
  ) |>
  visTable(groupColumn = "codelist_with_details_name")
```

:::{custom-style="Caption"}
**Table 1:** Demographics characterisation at first record.
:::

```{r}
arrangeCharacteristics <- function(x) {
  x |>
    mutate(order_id = dplyr::case_when(
      variable_name == "Number records" ~ 1L,
      variable_name == "Number subjects" ~ 2L,
      variable_name == "Sex" ~ 5L,
      variable_name == "Age" ~ 6L,
      variable_name == "Age group" ~ 7L,
      variable_name == "Ethnicity" ~ 8L,
      variable_name == "Socio-economic status" ~ 9L,
      variable_name == "MI type" ~ 12L,
      variable_name == "Prior comorbidities (-inf to 0]" ~ 13L
    )) |>
    arrange(order_id, variable_level) |>
    select(!"order_id")
}
characteristics <- data$summarise_demographics |>
  updateCohortName("group_level") |>
  filter(
    variable_name %in% c(
      "Number subjects", "Sex", "Age", "Age group",
      "Ethnicity", "MI type", "Socio-economic status", "Prior comorbidities (-inf to 0]"
    ),
    !(
      variable_level %in% c("Male", "Both") |
        grepl("cute mi", variable_level) |
        grepl("Stroke", variable_level) |
        grepl("docardit", variable_level) |
        grepl("arotid", variable_level)
      ) | 
      is.na(variable_level)
      ,
    !estimate_name %in% c("median", "q25", "q75", "min", "max"),
    strata_level == "overall"
  ) |>
  arrange(group_level, cdm_name) |>
  arrangeCharacteristics() |>
  mutate(
    variable_level = coalesce(variable_level, variable_name),
    variable_name = case_when(
      variable_name == "Age group" ~ "Age group — N(%)",
      variable_name == "Ethnicity" ~ "Ethnicity — N(%)",
      variable_name == "Socio-economic status" ~ "Socio-economic status — N(%)",
      variable_name == "MI type" ~ "STEMI/NSTEMI — N(%)",
      variable_name == "Prior comorbidities (-inf to 0]" ~ "Prior comorbidities (-inf to 0] — N(%)", 
      .default = "Demographics"
    ),
    variable_level = if_else(
      variable_name == "Demographics",
      case_when(
        variable_level == "Number records" ~ "Number records — N",
        variable_level == "Number subjects" ~ "Number subjects — N",
        variable_level == "Female" ~ "Female — N (%)",
        variable_level == "Age" ~ "Age — Mean (SD)",
        .default = variable_level
      ),
      case_when(
        variable_level == "Not stemi" ~ "NSTEMI",
        variable_level == "Stemi" ~ "STEMI",
        .default = variable_level
      )
    )
  )

characteristics |>
  filter(!grepl("ST-", group_level)) |>
  tableCharacteristics(
    hide = c("cohort_name", "strata", "table_name", "age_range", "ses", "sex", "mi_type", "estimate_name", "index_condition", "drug", "estimate_name"),
    header = c("cdm_name"),
    groupColumn = "variable_name"
  )
```

:::{custom-style="Caption"}
**Table S1:** Demographics characterisation at first record.
:::

```{r}
characteristics |>
  filter(grepl("ST-", group_level)) |>
  tableCharacteristics(
    hide = c("strata", "table_name", "age_range", "ses", "sex", "mi_type", "estimate_name", "index_condition", "drug", "estimate_name"),
    header = c("cohort_name", "cdm_name"),
    groupColumn = "variable_name"
  )
```


:::{custom-style="Caption"}
**Table 2:** Medications.
:::

```{r}
data$summarise_treatments |>
  updateCohortName("group_level") |>
  mutate(
    variable_level = str_to_title(variable_level),
    variable_level = if_else(
      variable_level == "Thrombolytics", "Thrombolytics*", variable_level
    )
  ) |>
  filter(strata_level == "overall") |>
  arrange(group_level, cdm_name) |>
  arrangeCharacteristics() |>
  CohortCharacteristics::tableCharacteristics(
    hide = c("variable_name", "strata", "table_name", "age_range", "ses", "sex", "mi_type", "estimate_name", "drug", "index_condition"),
    header = c("cdm_name"),
    groupColumn = c("cohort_name")
  )
```

:::{custom-style="Caption"}
**Figure 1:** 28-day mortality after Index date stratified by Age group, Ethnicity, Socio-Economic Status (SES) and Sex. Each colour represents a different hospital data.
:::

```{r, fig.height=11/2.54}
x <- data$summarise_death |>
  tidy() |>
  updateCohortName("cohort_name") |>
  mutate(
    strata_name = str_extract(strata, "^[^:]+"),
    strata_value  = coalesce(trimws(str_extract(strata, "(?<=:).*")), "")
  ) |>
  filter(strata != "overall")
p <- ggplot(data = x, mapping = aes(x = strata_value, y = percentage, fill = cdm_name)) +
  geom_col(position = "dodge") +
  facet_grid(. ~ strata_name, scales = "free_x") +
  labs(fill = "", y = "Percentage (%)", x = "") +
  theme_bw()
stylePlot(p) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 60, hjust = 1)
  )
```

:::{custom-style="Caption"}
**Figure 2:** Radial plots.
:::

```{r, fig.width=1.5*20/2.54, fig.height=1.5*15/2.54}
x <- data$radial |>
  updateCohortName("index_condition") |>
  filter(
    !variable_level %in% c("Mix", "Other", "Unknown", "Missing"),
    !grepl("ST", index_condition),
    !drug %in% "thrombolytics"
  ) |>
  mutate(
    value = count1 / den1,
    drug = clean(drug),
    variable_level = case_when(
      grepl("Q1", variable_level) ~ "Q1",
      grepl("Q5", variable_level) ~ "Q5",
      .default = variable_level
    )
  )
labs <- x |>
  distinct(variable_level, y) |>
  arrange(y) |>
  mutate(y_new = as.integer(row_number()))
x <- x |>
  left_join(labs, by = "y")
ggplot(
  data = x,
  mapping = aes(x = y_new, y = value, fill = variable_name)
) +
  geom_col() +
  facet_grid(cdm_name ~ drug) +
  coord_polar() +
  scale_x_continuous(
    breaks = labs$y_new,
    labels = labs$variable_level
  ) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  themeVisOmop(style = "_brand.yml") +
  theme(
    legend.position = "top",
    strip.text.y.right = element_text(angle = 270)
  ) +
  labs(x = "", y = "Percentage (%)", fill = "")
```

:::{custom-style="Caption"}
**Table 3:** Initiators OR.
:::

```{r}
#| results: asis
x <- data$drug_initiate |>
  updateCohortName("group_level") |>
  mutate(strata_level = clean(strata_level)) |>
  filter(
    group_level %in% c("Acute Myocardial Infarction"),
    strata_level %in% drugsTidy
  )

xp <- x |> 
  tidy() |>
  filter(drug != "Thrombolytics")
ref <- tribble(
  ~variable_name, ~variable_level, ~rr, ~rr_lower, ~rr_upper,
  "Age group", "60 to 69", 1, NA, NA,
  "Sex", "Female", 1, NA, NA,
  "Socio-economic Status", "1", 1, NA, NA,
  "Ethnicity", "White", 1, NA, NA,
  "MI type", "NSTEMI", 1, NA, NA,
) |>
  inner_join(
    xp |>
      distinct(cdm_name, index_condition, drug, model, model_type, variable_name),
    by = "variable_name"
  )
lab <- tribble(
  ~variable_name, ~variable_level,
  "Age group", "18 to 39",
  "Age group", "40 to 49",
  "Age group", "50 to 59",
  "Age group", "60 to 69",
  "Age group", "70 to 79",
  "Age group", "80 to 89",
  "Age group", "90 or above",
  "Sex", "Female",
  "Sex", "Male",
  "Socio-economic Status", "Q1 (Least dep.)",
  "Socio-economic Status", "Q2",
  "Socio-economic Status", "Q3",
  "Socio-economic Status", "Q4",
  "Socio-economic Status", "Q5 (Most dep.)",
  "Socio-economic Status", "Missing",
  "Ethnicity", "White",
  "Ethnicity", "Asian",
  "Ethnicity", "Black",
  "Ethnicity", "Mix",
  "Ethnicity", "Other",
  "Ethnicity", "Unknown",
  "MI type", "NSTEMI",
  "MI type", "STEMI",
  #"MI type", "Both",
  "MI type", "None"
) |>
  mutate(id = row_number())
xp <- xp |>
  union_all(ref) |>
  mutate(
    variable_level = case_when(
      variable_name == "Socio-economic Status" & variable_level == "1" ~ "Q1 (Least dep.)",
      variable_name == "Socio-economic Status" & variable_level == "5" ~ "Q5 (Most dep.)",
      variable_name == "Socio-economic Status" & variable_level %in% c("2", "3", "4") ~ paste0("Q", variable_level),
      .default = variable_level
    ),
    type = if_else(is.na(rr_lower), "reference", "estimate")
  ) |>
  inner_join(lab, by = c("variable_name", "variable_level")) |>
  arrange(drug, id)

xp <- xp |>
  filter(model_type == "Individual model") |>
  union_all(
    xp |>
      filter(model_type == "Combined model")
  )

for (dr in drugsTidy) {
  cat(paste0(dr, "\n\n"))
  tbl <- xp |>
    filter(drug == dr) |>
    mutate(or = if_else(
      is.na(rr_lower),
      "Ref.",
      sprintf("%.2f [%.2f - %.2f]", rr, rr_lower, rr_upper)
    )) |>
    select(!c("index_condition", "model_type", "drug", "rr", "rr_lower", "rr_upper", "type", "id")) |>
    pivot_wider(names_from = "cdm_name", values_from = "or") |>
    visTable(
      groupColumn = c("model"),
      hide = c()
    )
  cat(knitr::knit_print(tbl))
  cat("\n\nhi\n\n")
}
```

:::{custom-style="Caption"}
**Figure 3:** Initiators OR.
:::

```{r, fig.width=2*20/2.54, fig.height=2*15/2.54}
xp <-  xp |> 
  filter(model_type == "Individual model")
ggplot(
  data = xp,
  mapping = aes(
    x = rr,
    y = -id,
    xmin = rr_lower,
    xmax = rr_upper,
    colour = variable_name
  )
) +
  geom_vline(xintercept = 1, colour = "black") +
  geom_errorbar() +
  geom_point(mapping = aes(shape = type)) +
  scale_shape_manual(
    values = c(
      reference = 15,  # square
      estimate = 16    # circle
    ),
    guide = "none"
  ) +
  facet_grid(cdm_name ~ drug) +
  scale_x_log10(
    name = "Odds Ratio",
    breaks = c(0.125, 0.25, 0.5, 1, 2, 4, 8),
    labels = c("1/8", "1/4", "1/2", "1", "2", "4", "8")
  ) +
  coord_cartesian(xlim = c(0.1, 10)) +
  scale_y_continuous(
    breaks = -lab$id,
    labels = lab$variable_level,
    name = ""
  ) +
  themeVisOmop("_brand.yml") +
  theme(
    strip.text.y.right = element_text(angle = 270),
    legend.position = "top"
  ) +
  labs(colour = "")
```

```{r}
data$summarise_procedures |>
  updateCohortName("group_level") |>
  filter(
    group_level %in% c("Acute Myocardial Infarction"),
    cdm_name %in% c("Barts", "LTHT")
  ) |>
  union_all(
    data$summarise_procedures |>
      updateCohortName("group_level") |>
      filter(
        group_level %in% c("Acute Myocardial Infarction ST-Elevation Myocardial Infarction"),
        cdm_name %in% c("Barts", "LTHT"),
        strata_level == "overall"
      ) |>
      mutate(
        group_level = "Acute Myocardial Infarction",
        strata_name = "strata",
        strata_level = "STEMI"
      )
  ) |>
  union_all(
    data$summarise_procedures |>
      updateCohortName("group_level") |>
      filter(
        group_level %in% c("Acute Myocardial Infarction Non-ST-elevation Myocardial Infarction"),
        cdm_name %in% c("Barts", "LTHT"),
        strata_level == "overall"
      ) |>
      mutate(
        group_level = "Acute Myocardial Infarction",
        strata_name = "strata",
        strata_level = "NSTEMI"
      )
  ) |>
  visOmopTable(
    estimateName = c("N (%)" = "<count> (<percentage>%)"),
    header = "cdm_name",
    hide = c("cohort_name", "index_condition", "drug", "variable_name", "estimate_name"),
    groupColumn = "variable_level"
  )
```


